### 2014-11-05 D1-3.tex
Added Privacy Policy Section
WIP

\documentclass[external]{20120615_deliverable_template_ukob}

\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
%\usepackage{showkeys}

@@

\LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
{}

@@

\chapter{Introduction}
\label{chap:Introduction}

@@

\chapter{Implementation of Privacy Protection Measures}
\subsection{Privacy Policy}
\subsection{Privacy Dashboard}
\subsection{Transfer Encryption (HTTPS/SSL)}
\subsection{Database Access Policies}
\subsection{Anonymization of recorded trips}
\subsection{Randomization of GPS data}

>>>>

\documentclass[external]{20120615_deliverable_template_ukob}
\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{listings}
\usepackage{subfigure}
%\usepackage{showkeys}

@@

\LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
{\small maxmeffert@uni-koblenz.de}

@@

\chapter{Introduction}\label{chap:Introduction}
In recent days the importance of privacy protection has been amplified
by the reports about the mass surveillance of ordinary citizens on a
global scale by the NSA and other intelligence agencies around the
world.

While aiming at the noble cause of enhancing eParticipation using
mobile technologies, Live+Gov systems do process a large variety data
that is potentially infringing the citizens privacy. The captured data
includes personal information like name, phone numbers and email
addresses and sensor data from GPS and accelerometer sensors. Also
with some applications it is possible to gather images and textual
input from the citizen.

While the collection of this data is necessary for providing the
advanced services that Live+Gov aims to deliver, at he same time, the
available raw data can be used to draw a very detailed picture of the
private life of the citizen. For instance can GPS location tracking be
used to reveal shopping habits (e.g. when a car seller is visited) and
associations to political groups (when a meeting is
attended). Accelerometer data can be used to infer medical conditions
like walking disabilities.

Images can contain faces of nearby persons to with whom the citizen is
associated. All this data is highly sensitive to the citizens privacy
and can be used against the citizen if it falls in the wrong hands.

The great importance of protecting the citizens privacy should be
apparent from these examples. The European Union, as well as many
other countries in the past, has set out a number of directives that
regulate the collection, processing and use of privacy sensitive
data. The most important legislation on the European Level is the Data
Protection Directive 95/46/EC [1], which is also summarized in the
European Charta of Fundamental Rights [2] of EU citizens:

\begin{quote}
Article 7. Respect for private and family life.

Everyone has the right to respect for his or her private and family
life, home and communications.

Article 8. Protection of personal data.

(1) Everyone has the right to the protection of personal data
concerning him or her.

(2) Such data must be processed fairly for specified purposes and on
the basis of the consent of the person concerned or some other
legitimate basis laid down by law.  Everyone has the right of access
to data which has been collected concerning him or her, and the right
to have it rectified.

(3) Compliance with these rules shall be subject to control by an
independent authority.  The Live+Gov fully consortium complies with
this legislation and is moreover committed to convey and enact the
ethical values inherent in these texts.
\end{quote}

Privacy as Control over Private Data

The ethical aspects of privacy have been the subject of study of many
social scientists and philosophers [3]. Many researchers do focus on
the ways in which privacy can be infringed. Thus they invest a great
amount of work in exploring threats, instead of describing why privacy
is so valuable to us. One scholar which follows an alternative aproach
in the context of digital monitoring is Charles Fried [4]. He
investiages, why we are intuitively so sensitive to violations of our
privacy. For him privacy is not asserted as an intrinsic value by
itself, he rather sated:

\begin{quote}
Privacy is not simply an absence of information about us in the minds of others;
rather it is the control we have over information about ourselves.
\end{quote}

Depending on the conversation partner, we change the degree of
information we share if it is a total stranger, a colleague, close
friend or a doctor. With doctors we share information of great
intimacy we do not share with anyone else.

Moreover, we trust those friends to not reveal information about us to
others by respecting their privacy. Trust needs the possibility of
unknown failure. If we would constantly monitor our partners, they
cannot fail unnoticed nor can they willingly share that information
with us.  So privacy, according to Fried, is the foundation of our
core relations. And thus it is valuable, because those relations are
essential to human society.

Fried’s study on the understanding of privacy provided a great
contribution to the research on the same term in philosophy and
computer science [5] and despite the fact his text was published in
1970, he already included technologies to its viewpoint (like location
monitoring) that are particularly relevant to our Context.

Live+Gov Privacy Dashboard gives citizens back the Control

In order to meet the the legal requirements and offer the citiens to
keep a maximum of contol about their personal data, Live+Gov systems
implements a number of privacy protection measure. One of the measures
that is under current development is a Privacy Dashboard (Figure
1). Using this dashboard the citizen will be able to take the
following actions: View and export all data from the citizen that is
currently stored in the Live+Gov system Selectively delete parts of
the stored data Get information about processing applied to this data
Consent or decline selectively to processing steps

View the end poducts of data mining applied to his data in a
comprehensive, graphical way These measures provide the citizen with a
maximal respect for their privacy while at the same time allowing
advanced data processing. Although, abusive exploitation of the data
is still possible, the level of transparency offered by the dashboard
builds trust between the citizen and the service provider.

@@

\input{Implementations}

\chapter{Summary}\label{chap:Summary}

### 2014-11-05 Implementations.tex

\chapter{Implementation of Privacy Protection Measures}
\section{Privacy Policy}

@@

\section{Privacy Dashboard}
The Live+Gov Privacy Dashboard is a software component that allows the
citizens to view the data about themselves that is stored in the
Sensor Data Storage server.  The foundation of the Privacy Dashboard
component is the inspection tool that was introduced in D1.1 as a
means to perform basic quality controlling of the collected data.

The Privacy Dashboard is currently endowed with the following views:
\begin{itemize}
  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
    access to the dashboard is granted, the user is prompted for
    credentials.  The credentials are generated by the sensor
    collection application on the mobile device.  The username can be
    freely chosen by the user. The password is then generated by the
    mobile application and transferred to the server alongside with
    the recording.
 \item \textbf{Recording Overview.} (Cf. Figure \ref{fig:PDOverview}).
   This view shows all recordings that the user submitted to the Data
   Storage Service including basic meta information like start data of
   recording, duration. The user can click on each recording in order
   to view the recorded data in the Raw Data View.

@@

 \item \textbf{Raw Data View.} (cf. Figure \ref{fig:PDRawData}) The
   raw data view shows all recorded raw data to the user.  GPS data is
   visualized on a map. Motion sensor data is displayed as plots. Wifi
   and GSM data is displayed in lists.
   The raw data from each sensor can be downloaded individually as CSV file.

 \item \textbf{Processing View.} (cf. Figure
   \ref{fig:PDProcessedData}).  The processing view shows the data
   mining end-products for each processing step.  The Figure shows the
   processing view for the Activity Recognition. On the top a
   time-line is show that visualizes the different recognized
   activities via color-coding. A map below displays the corresponding
   GPS track with the same color coding.

   A similar view for Service Line Detection results is currently
   under development.
\end{itemize}

@@

@@

\section{Data Expiring Dates}
Trips stored in the Sensor Data Storage Service have been endowed with
a data expiration date. A python script is provided that performs the
of expired data. In our installation we have this script running as a
cron job every night. The resulting schema of the trip table listed in
Figure \ref{fig:TripScheme}.

@@

\begin{figure}
\centering
\begin{verbatim}
  Column  |          Type
----------+------------------------
 trip_id  | integer
 user_id  | character varying(36)
 start_ts | bigint
 stop_ts  | bigint
 name     | character varying(255)
 deleted  | boolean
 expires  | bigint
\end{verbatim}
\caption{Database Schema of the Trip Table}
\label{fig:TripScheme}
\end{figure}

@@

anonymization: Sets all user_ids to 0
    Option:
        SQL Select Query: Every row found by this query will
        be set to user_id = 0.

@@

blur: Offsets every GPS point by a random amount
    Options:
        The maximum distance a point can have to its origin

k_anonymity: Creates a centroid for every GPS point
    Options:
        First - Number of maximum GPS points to use. If 0 then
                we grab every point inside the radius
        Second - Maximum distance a point can have to be used
                for the creation of the new point
\end{verbatim}
\caption{Usage options for the privacy.py script.}
\end{figure}

\section{K-Anonymity of GPS routes}
Python script that deletes all parts of journeys that have are not
near to at least k other journeys.

TODO: Insert Literature Reference.
TODO: Add screenshots.

\section{Randomization of GPS data}
Python Script that adds random noise to GPS data.

>>>>

\chapter{Implementation of Privacy Protection Measures}
\section{Privacy Policy} \label{sec:PP}

@@

\section{Privacy Dashboard}
The Live+Gov Privacy Dashboard is a software component that allows the
citizens to view and control the data about themselves that is stored
in the Sensor Data Storage server.  The foundation of the Privacy
Dashboard component is the inspection tool that was introduced in D1.1
as a means to perform basic quality controlling of the collected data.

The Privacy Dashboard is currently endowed with the following views:
\begin{itemize}
  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
    access to the dashboard is granted, the user is prompted for
    credentials.  The credentials are generated by the sensor
    collection application on the mobile device.  The username can be
    freely chosen by the user. The password is then generated by the
    mobile application and transferred to the server alongside with
    the recording.

 \item \textbf{Recording Overview} (Cf. Figure \ref{fig:PDOverview}).
   This view shows all recordings that the user submitted to the Data
   Storage Service including basic meta information like start data of
   recording, duration. The user can click on each recording in order
   to view the recorded data in the Raw Data View.

@@

 \item \textbf{Raw Data View} (cf. Figure \ref{fig:PDRawData}).
   The raw data view shows all recorded raw data to the user.  GPS
   data is visualized on a map. Motion sensor data is displayed as
   plots. Wifi and GSM data is displayed in lists.  The raw data from
   each sensor can be downloaded individually as CSV file.

 \item \textbf{Processing View} (cf. Figure \ref{fig:PDProcessedData}).
   The processing view shows the data mining end-products for each
   processing step.  The Figure shows the processing view for the
   Activity Recognition. On the top a time-line is show that
   visualizes the different recognized activities via color-coding. A
   map below displays the corresponding GPS track with the same color
   coding.

   A similar view for Service Line Detection results is currently
   under development.
 \item \textbf{Privacy Settings.}
   This view contains the privacy policy shown in Section \ref{sec:PP}.

   We intend to include settings also settings concerning the
   expiration date of the recordings, agreements to certain processing
   steps, view logs of performed processing.
\end{itemize}

The Privacy Dashboard is implemented on a
nodes.js\footnote{\url{http://nodejs.org}} web server using
Express.js\footnote{\url{http://expressjs.com/}} as a web-framework
and d3.js\footnote{\url{http://d3js.org/}} as plotting library. The
data is pulled directly from the PostgreSQL database of the Data
Storage Service.

@@

\pagebreak

@@

\section{Data Expiring Dates}
Trips stored in the Sensor Data Storage Service have been endowed with
a data expiration date. A python script is provided that performs the
of expired data. In our installation we have this script running as a
cron job every night. The resulting schema of the trip table listed in
Figure \ref{fig:TripScheme}. The expiring date is encoded as a 64bit
integer holding the unix timestamp in seconds of the expiration date.

@@

\begin{figure}
\begin{center}
\begin{alltt}
  Column  |          Type
----------+------------------------
 trip_id  | integer
 user_id  | character varying(36)
 start_ts | bigint
 stop_ts  | bigint
 name     | character varying(255)
 deleted  | boolean
 \textbf{expires}  | \textbf{bigint}
\end{alltt}
\end{center}
\caption{Database Schema of the Trip Table}
\label{fig:TripScheme}
\end{figure}

@@

anonymization: Randomizes user_ids
    Option:
        SQL Select Query: Every row found by this query will
        be randomized.

@@

blur: Offsets every GPS point by a random amount
    Options:
        The expected distance a point has to its origin

k_anonymity: Creates a centroid for every GPS point
    Options:
        First - Number of maximum GPS points to use. If 0 then
                we grab every point inside the radius
        Second - Maximum distance a point can have to be used
                for the creation of the new point
\end{verbatim}
\caption{Usage options for the privacy.py script.}
\label{fig:privacypy}
\end{figure}

\section{GPS Anonymization: Gaussian Noise}

The Gaussian Noise anonymization method adds random noise to the
stored GPS points. The \texttt{privacy.py} script (cf. Figure
\ref{fig:privacypy}) implements this feature as "blur'' and allows the
specification of the displacement variance.  Figure \ref{fig:noise}
shows an example of a GPS track with and without added noise.

The Gaussian Noise anonymization is a basic method is usually mixed
well with other methods.

\begin{figure}
  \centering
  \subfigure[Original GPS samples]{
    \includegraphics[width=0.45\textwidth]{screenshots/GPS_sample.png}
    }
  \quad
  \subfigure[GPS samples with noise]{
    \includegraphics[width=0.45\textwidth]{screenshots/AfterBlur.png}
  }
  \caption{GPS Anonymization with Gaussian Noise}
  \label{fig:noise}
\end{figure}

\section{GPS Anonymization: K-Anonymity}
This method of GPS anonymization is a variant of Sweeney's K-Anonymity
\cite{sweeney2002k} applied to GPS data along the lines of Grutser and
Grunwald \cite{Gruteser2003}.  The basic idea is that the resulting
data from a single person should not distinguishable form data from
K-1 other persons. In order to achieve such kind of anonymization we
calculate for each GPS point $P$ by a user $U$ the $K-1$ nearest $GPS$
points $Q_1, \dots, Q_{K-1}$ which are recorded by $K-1$ different
users.  The anonymized point $R$ is now the centroid of the $K$ points
$(P,Q_1,\dots,Q_{K-1})$. This operations applied to all data points
$P$ in the dataset. The resulting anonymized points $R$ are stored in
a new table, so that these points do not take part in further the
centroid calculations.

Figure \ref{fig:k-ano} shows the results of application of K-anonymity
to the original GPS trip displayed in Figure \ref{fig:noise}.
$K$-Anonymity implementation script (cf. Figure \ref{fig:privacypy})
supports a second parameter that specifies the maximal distance that a
point can get displaced by this method.

\begin{figure}
  \centering
  \subfigure[K=7 anonymization]{
    \includegraphics[width=0.4\textwidth]{screenshots/k_anon_7_nearest_points.png}
  }
  \quad
  \subfigure[K=15 Anonymization]{
    \includegraphics[width=0.4\textwidth]{screenshots/k_ano_15_nearest_points.png}
  }
  \caption{GPS Anonymization with K-Anonimity}
  \label{fig:k-ano}
\end{figure}

\section{GPS Anonymization: Haircut}

The \emph{haircut} anonymization method was developed by UKob to
address the problem of GPS tracks that show locations, like the home
address, that are unique to a single user. Looking on our data we see
many trips that originate at a unique location and later on meet other
tracks at locations of public interest like bus stops or larger
streets. The haircut methods cuts away parts of tracks that
"stick-out'' the whole set of recorded trajectories. For a given GPS
location $P$ the number of users is computed that have GPS samples
near to $P$. If this number is lower than a given threshold (e.g. $3$)
the point $P$ is dropped from the dataset.

The implementation script (cf. Figure \ref{fig:privacypy}) takes two
supports two parameters that specify the radius to look for other GPS
samples and the number of different users that are required to have
samples inside this radius. Figure \ref{fig:haircut} shows an example
a two-trips before and after the haircut.

\begin{figure}
  \centering
  \subfigure[Original GPS tracks]{
    \includegraphics[width=0.4\textwidth]{screenshots/haircut_pre.png}
  }
  \quad
  \subfigure[GPS tracks with haircut]{
    \includegraphics[width=0.4\textwidth]{screenshots/haircut_post.png}
  }
  \caption{GPS Anonymization with Haircut Method}
  \label{fig:haircut}
\end{figure}
